---
title: "Predicting Delayed Flights"
author: "Sonal Kanabar"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Project Background
*Outcome of interest : whether or not the flight is delayed (delayed here means arrived more than 15 minutes late).

*% of delayed flights among 2201 flights = 19.5%. 

*The data were obtained from the Bureau of Transportation Statistics (available on the web at www.transtats.bts.gov). 

*GOAL :-)  to accurately predict whether or not a new flight (not in this dataset), will be delayed.

*The outcome variable is whether the flight was delayed, and thus it has two classes (1 = delayed and 0 = on time).


####Packages Used 
```{r}
library(caret)   #for creating data partition
library(e1071)   #for creating Naive Bayes
library(gmodels) #for creatong cross table
library(pROC)    #for plotting ROC curves

```

####Data Exploration
```{r}
#Reading the dataset in R
read_data <- read.csv("FlightDelays.csv")

#Subsetting the data by selecting only the predictors that is been asked to use
Flight_data <- subset(read_data, select= c(DAY_WEEK, ORIGIN, DEST, CARRIER, CRS_DEP_TIME, DEP_TIME, Flight.Status))

summary(Flight_data)
#After running the summary statistics we figured that we need to convert Week and Time variables into factors.
```

####Data Transformation
```{r}
#Week and Time variables need to recoded as factors
names<- c('DAY_WEEK' , 'CRS_DEP_TIME', 'DEP_TIME')
Flight_data[,names] <- lapply(Flight_data[,names], factor)
summary(Flight_data)
```

####Data Partition
```{r}
#Dividing the data into 60% training and 40% validation
set.seed(2019)
#library(caret)
#Divide data into validation and training
Index_Train<-createDataPartition(Flight_data$CARRIER, p=0.6, list=FALSE) #Train Data
Train_data <-Flight_data[Index_Train,] #1323 rows
Test_data  <-Flight_data[-Index_Train,] #Validation Data(remaining) #878

#NROW(Train) (number of rows in training data = 1323)
#NROW(Test)  (number of rows in test data     = 878 )
```

####Training the data
```{r}
#Building Naive Bayes model 
library(e1071)
NB_training_model <- naiveBayes(Flight.Status ~. , data= Train_data )
NB_training_model
#First it shows the ratios of default (yes) and default (no) in the training set (called a priori probabilities)
#Then it will give table for each target class, mean and standard deviation of the (sub-)variable. 
```

####Testing the data
```{r}
#Now, use the model on the test set

#Predicting the default status of test dataset 
Predicted_Test_labels <-predict(NB_training_model,Test_data)

#library("gmodels")

#Show the confusion matrix of the classifier
CrossTable(x=Test_data$Flight.Status,y=Predicted_Test_labels, prop.chisq = FALSE) 
```


```{r}
#Raw prediction probabilities rather than the predicted class

nb_model <- naiveBayes(Flight.Status ~. , data= Train_data )

#Make predictions and return probability of each class
Predicted_Test_labels <-predict(NB_training_model,Test_data, type = "raw")

#show the first few values 
head(Predicted_Test_labels)
```

####ROC Curves
```{r}
# install.packages("pROC") # install if necessary
library(pROC)
#Passing the first column of the predicted probabilities 
#That column contains the probability associate to 'delayed'
roc(Test_data$Flight.Status, Predicted_Test_labels[,1]) #0.6602
plot.roc(Test_data$Flight.Status, Predicted_Test_labels[,1])
```

####Confusion Matrix
```{r}
Predicted_Test_labels <-predict(NB_training_model,Test_data)
Actual <- Test_data$Flight.Status
table(Predicted_Test_labels,Actual)
```

